---
title: "Problems Sheet 3"
subtitle: "Statistics 1"
author: "Dom Hutchinson"
output:
  pdf_document: 
     fig_width: 6
     fig_height: 4
  html_notebook: default
  html_document: default
  word_document: default
---
# Part A

## Question 2

### a)

```{r}
xvalues<-rexp(10000,rate=1)
xsamples<-matrix(xvalues,nrow=1000)
```
### b)

```{r}
sample.mean<-apply(xsamples,1,mean)
theta.mle<-1/sample.mean
hist(theta.mle)
```

### c)
```{r}
sample.median=apply(xsamples,1,median)
tau.mle<-log(2)/theta.mle
boxplot(sample.median,tau.mle, names=c("Sample Median","MLE"))
abline(h=log(2))
```

### d)
Both calculations produce medians which are practicaly the same as the actual median.
The sample median calculation has a much larger range & significantly more outliers than that of the maximum likelihood estimate.

### e)
```{r}
sample.mean<-mean(sample.median)
sample.var <-var(sample.median)
mle.mean   <-mean(tau.mle)
mle.var    <-var(tau.mle)

sample.mean
sample.var
mle.mean
mle.var

sample.bias=sample.mean-log(2)
mle.bias   =mle.mean-log(2)
sample.bias
mle.bias

sample.mse=sample.var+sample.bias^2
mle.mse=mle.var+mle.bias^2
sample.mse
mle.mse
log(2)
```

Both methods have a positive bias in their results, however maximum likelihood estimate has a much lower value. Maximum likelihood estimate has a lower variance & mean-square error