---
title: "Problems Sheet 5"
subtitle: "Statistics 1"
author: "Dom Hutchinson"
output:
  pdf_document: 
     fig_width: 6
     fig_height: 4
  html_notebook: default
  html_document: default
  word_document: default
---
```{r}
load(url("https://people.maths.bris.ac.uk/~maxca/stats1/stats1.RData"))
```

# Part A
## Question 1
**a)** Let $\mu_N$ be the mean $N_2O$ in the new engines \& $\mu_O$ be the $N_2O$ in the old engines.  
Define hypothesis test $H_0:\mu_N=\mu_O$ \& $H_1:\mu_N<\mu_O$.  
Since both populations have the same variance a pooled estimate t-test is appropriate.  

**b)** Let $\mu_M$ be the mean systolic blood pressure in men \& $\mu_W$ be the mean systolic blood pressure in women.  
Define hypothesis test $H_0:\mu_M=\mu_W$ \& $H_1:\mu_M\neq\mu_W$.  
Since the two samples are independent \& variances are unknown a two-sample t-test is appropriate.  

**c)** Let $\mu_A$ be the mean reaction time after an alcholic drink \& $\mu_B$ be the mean reaction before an alcoholic drink.  
Define hypothesis tests $H_0:\mu_A=\mu_B$ \& $H_1:\mu_A<\mu_B$.  
Since the samples are paired \& the variances are unknown a two-sample t-test is appropritate.  

**d)** Let $\mu$ be the mean weight of a chocolate bar.  
Define hypothesis test $H_0:\mu=50$ \& $H_1:\mu<50$.  
Since there is only one sample \& the variance is unknown a one-sample t-test is approperiate.  

## Question 4
Let $S_i$ denote the time of runner $i$ at sea-level \& $A_i$ denote the time of runner $i$ at high altitude. 
Define $D_i:=A_i-S_i$ to be how much slower runner $i$ was at sea-level compared to high altitude.  
We assume that $W_1,\dots,W_8$ is a simple random sample from $N(\mu_w,\sigma_w^2)$ distribution with $\mu_w$ \& $\sigma_w^2$ unknown.  
Define hypothesis test $H_0:\mu_w=0$ \& $H_1:\mu_w>0$.  
```{r}
W=runner$high-runner$sea # Sample of time differences
WBar<-mean(W)            # Sample mean
S2_W<-var(W)             # Sample variance
cat("WBar=",WBar,", S2_W=",S2_W,sep="")
```
Define test statistic $T(W_1,\dots,W_8)=\sqrt{n}\dfrac{\bar{W}}{\sigma_w}\sim t_7$.  
Thus
```{r}
t_obs<-sqrt(length(W))*(WBar/sqrt(S2_W)) # Observed test statistic
cat("t_obs=",t_obs,sep="")
```
Thus $p=\mathbb{P}(T>t_{obs}|H_0\ \mathrm{true})=\mathbb{P}(t_7>2.195823)=1-\mathbb{P}(t_7<2.195823)$
```{r}
cat("p = 1-",pt(t_obs,7)," = ",1-pt(t_obs,7),sep="") # p-value
```
Since $p=0.0321<0.05=\alpha$ there is strong evidence to reject $H_0$ in favor of $H_1$.  
Thus I conclude that running at high altidude did have an affect on the runners' times, when compared to their times at sea-level.  

# Part B
## Question 4
**a)**
```{r}
X=rain$spring # Define samples
Y=rain$autumn

ss_xy<-sum(X*Y)-length(X)*mean(X)*mean(Y)
ss_xx<-sum(X^2)-length(X)*(mean(X)^2)

bHat<-ss_xy/ss_xx
aHat<-mean(Y)-bHat*mean(X)

cat("ss_xy=",ss_xy,", ss_xx=",ss_xx,", aHat=",aHat,", bHat=",bHat,sep="")
plot(X,Y)
abline(aHat,bHat)
```

**b)**
```{r}
fitted<-aHat+bHat*X
residuals<-Y-fitted
m<-matrix(c(round(fitted,2),round(residuals,2)),10,2)
colnames(m)<-c("fitted","residuals")
rownames(m)<-(1:10)
t(m)
```

```{r}
plot(X,residuals,xlab="Spring Rainfall")
segments(X,0,X,residuals)
abline(0,0)
```

From the residuals plot there is little evidence of systematic error, so the linear regression model is sufficiently good to use.  

## Question 5
Consider a model where the response variable $Y$ is defined as $Y_i:=\gamma x_i+e_i$ where $x_i$ are values from the predictor variable \& $e_i$ are iid with $e_i\sim N(0,\sigma^2)$.  
For this model $\mathbb{E}(Y_i|x_i)=\gamma x_i$ so the least squares estimate is the value that minimises $\sum_{i=1}^n(y_i-\gamma x_i)^2$.  
These values can be found by setting the derivate of this sum to be 0.  
\[\begin{array}{rrcl}
&\frac{\partial}{\partial\gamma}\sum_{i=1}^n(y_i-\gamma x_i)^2&=&0\\
\implies&\frac{\partial}{\partial\gamma}\sum_{i=1}^ny_i^2-2\gamma x_iy_i+\gamma^2x_i^2&=&0\\
\implies&\sum_{i=1}^n-2x_iy_i+2\gamma x_i^2&=&0\\
\implies&\gamma\sum_{i=1}^nx_i^2&=&\sum_{i=1}^nx_iy_i\\
\implies&\hat{\gamma}&=&\dfrac{\sum_{i=1}^nx_iy_i}{\sum_{i=1}^nx_i^2}\\
\end{array}\]
We estimate $\sigma^2$ as $\hat{\sigma}^2=\frac{RSS}{n-2}$.  
\[\begin{array}{rrcl}
\mathrm{We\ have}&RSS&=&\sum_{i=1}^n\hat{e}_i^2\\
&&=&\sum_{i=1}^n(y_i-\hat{y}_i)^2\\
&&=&\sum_{i=1}^n(y_i-\hat{\gamma}x_i)^2\\
&&=&\sum_{i=1}^ny_i^2-2\hat{\gamma}x_iy_i+\hat{\gamma}^2x_i^2\\
&&=&\sum_{i=1}^ny_i^2-2\hat{\gamma}\sum_{i=1}^nx_iy_i+\hat{\gamma}^2\sum_{i=1}^nx_i^2\\
&&=&\sum_{i=1}^ny_i^2-2\dfrac{\left(\sum_{i=1}^nx_iy_i\right)^2}{\sum_{i=1}^nx_i^2}+\dfrac{\left(\sum_{i=1}^nx_iy_i\right)^2\sum_{i=1}^nx_i^2}{\left(\sum_{i=1}^nx_i^2\right)^2}\\
&&=&\sum_{i=1}^ny_i^2-\dfrac{\left(\sum_{i=1}^nx_iy_i\right)^2}{\sum_{i=1}^nx_i^2}\\
\implies&\hat{\sigma^2}&=&\frac{1}{n-2}\left(\sum_{i=1}^ny_i^2-\dfrac{\left(\sum_{i=1}^nx_iy_i\right)^2}{\sum_{i=1}^nx_i^2}\right)
\end{array}\]